AA: the Algorithmic Auto-regulation (Software Development) Methodology

### 2013-05-31 paper.tex
spell checked all paper and frontmatter

Dispite the advantages of GSD, it is known how
difficult it is to coordinate and fund free software on a larger scale
than currently available. That is, when teams are very heterogeneous, containing
not only volunteers and experienced developers, but also
contractors from different backgrounds and cultures. This observations
are in consonance with dedicated studies, such as~\cite{carmel1999}.
As a common base, related difficulties for GSD gather around distance, time and cultural differences. In
the case of free or open software projects, all these factors are
involved.

@@

To address these matters is the purpose of AA, a methodology and an associated software system
for coordinating distributed team work dealing with the disadvantages
of GSD. Team members take on an egalitarian role, and stay
voluntarily \textit{logged} in the system for part of their time
(e.g.\ 2 hours per day), during which they log a periodical short text
sentence --- similar to a ``tweet'' in this Twitter era --- as the
status of their activity using an easy to use command. These
``microblog sentences'' are publicly aggregated and validated by other
team members. Through AA, the community has a methodology and an associated
system to validate and enable the activities of a distributed software
team. It implicitly legitimizes financial support for the expansion of
the activity of the developer team as the activities are documented and validated.
The AA methodology is specially
useful for coordinating distributed and decentralized team work,
providing effective means to asynchronously update different team
members, diminishing the need for harsh synchronous meetings.

@@

\indraftnote{
A very good article on the value of asynchronous communication for personal
and group productivity, related to the key necessity of having moments of
introversion to avoid daily pressures of forced socialization. The way we work
on the digital age enables people to be very productive, the article also
mentions linux as a hallmark example~\cite{Thompson:Wired:2012}
}

@@

Various methodologies for GSD were built around the factors that
affect distributed team works, proposed by Carmel~\cite{carmel1999}
and comprising three distances: geographical, cultural and
temporal. Geographical distance prejudice \emph{coordination}, being
the act of integrating all the tasks distributed between
units~\cite{carmel2001}; \emph{control}, or the process to maintain
specific goals, policies or quality levels; and
\emph{communication}. All these factors are correlated, for example, a
team need to have clearly communication to work on tasks of a specific
problem. Cultural distance encompass differences on organizational and
natural culture. Spoken language, unit and ethnic values are common
forms of this distance. Some companies prefer to situate development
units in foreign locations with minimal cultural distance (e.g.\ an
American firm prefer Ireland, because of spoken language
similarity~\cite{carmel2001}). And finally the temporal distance that
prejudices synchronous communications like telephone or
videoconferences. Units of developers working on different time-zones
are concerned with managing of their agenda guided by this temporal
distance.

@@

\begin{table}
  \caption{Free and open software projects that received contributions
    from ``Macambiras''. On the first column we can see a list of
    applications of those projects. At right, the pseudonym of the
    ``Macambiras'' who sent \textit{commits} to the application. At
    ``Lab Macambira'', and at free software community in general, is a
    common practice to use pseudonym as identification.}
    \begin{tabular}{|l|l|}
        \hline
        Application           & ``Commiters''                       \\
        \hline \hline
        Mozilla Firefox       & daneoshiga, bzum                    \\
        Evince                & hick209, bzum, marcicano, mquasar   \\
        BePDF / Xpdf          & marcicano                           \\
        Ekiga                 & flecha                              \\
        Empathy               & fefo                                \\
        Lib Folks (Telepathy) & kamiarc                             \\
        Scilab                & v1z, humannoise                     \\
        VxL                   & v1z                                 \\
        ImageMagick           & v1z                                 \\
        OpenOffice            & hick209                             \\
        Puredata              & v1z, automata, greenkobold, gilson, bzum \\
        Puredata OpenCV       & v1z                                 \\
        Puredata GEM          & v1z, fefo, hick209                  \\
        Puredata PDP          & v1z, fefo, hick209                  \\
        ChucK                 & rfabbri, automata                   \\
        ChucK MiniAudicle     & rfabbri, automata                   \\
        WebRTC                & automata                            \\
        OSC-Web               & automata                            \\
        Web-PD-GUI            & automata                            \\
        Live-Processing       & automata                            \\
        ChucK-Wiimote         & automata                            \\
        Audiolet              & automata                            \\
        Extempore             & automata                            \\
        \hline
    \end{tabular}
    \label{tabela:contribuicoes}
\end{table}

>>>>

Despite the advantages of GSD, it is known how difficult it is to coordinate and
fund free software on a larger scale than currently available. That is, when
teams are very heterogeneous, containing not only volunteers and experienced
developers, but also contractors from different backgrounds and cultures. This
observations are in consonance with dedicated studies, such
as~\cite{carmel1999}.  As a common base, related difficulties for GSD gather
around distance, time and cultural differences. In the case of free or open
software projects, all these factors are involved.

@@

To address these matters is the purpose of AA, a methodology and an associated
software system for coordinating distributed team work dealing with the
disadvantages of GSD. Team members take on an egalitarian role, and stay
voluntarily \textit{logged} in the system for part of their time (e.g.\ 2 hours
per day), during which they log a periodical short text sentence --- similar to
a ``tweet'' in this Twitter era --- as the status of their activity using an
easy to use command. These ``micro-blog sentences'' are publicly aggregated and
validated by other team members. Through AA, the community has a methodology and
an associated system to validate and enable the activities of a distributed
software team. It implicitly legitimizes financial support for the expansion of
the activity of the developer team as the activities are documented and
validated.  The AA methodology is specially useful for coordinating distributed
and decentralized team work, providing effective means to asynchronously update
different team members, diminishing the need for harsh synchronous meetings.

@@

\indraftnote{
A very good article on the value of asynchronous communication for personal
and group productivity, related to the key necessity of having moments of
introversion to avoid daily pressures of forced socialization. The way we work
on the digital age enables people to be very productive, the article also
mentions Linux as a hallmark example~\cite{Thompson:Wired:2012}
}

@@

Various methodologies for GSD were built around the factors that affect
distributed team works, proposed by Carmel~\cite{carmel1999} and comprising
three distances: geographical, cultural and temporal. Geographical distance
prejudice \emph{coordination}, being the act of integrating all the tasks
distributed between units~\cite{carmel2001}; \emph{control}, or the process to
maintain specific goals, policies or quality levels; and
\emph{communication}. All these factors are correlated, for example, a team need
to have clearly communication to work on tasks of a specific problem. Cultural
distance encompass differences on organizational and natural culture. Spoken
language, unit and ethnic values are common forms of this distance. Some
companies prefer to situate development units in foreign locations with minimal
cultural distance (e.g.\ an American firm prefer Ireland, because of spoken
language similarity~\cite{carmel2001}). And finally the temporal distance that
prejudices synchronous communications like telephone or video-conferences. Units
of developers working on different time-zones are concerned with managing of
their agenda guided by this temporal distance.

@@

\begin{table}
  \caption{Free and open software projects that received contributions
    from ``Macambiras''. On the first column we can see a list of
    applications of those projects. At right, the pseudonym of the
    ``Macambiras'' who sent \textit{commits} to the application. At
    ``Lab Macambira'', and at free software community in general, is a
    common practice to use pseudonym as identification.}
    \small\begin{tabular}{|l|l|}
        \hline
        Application           & ``Commiters''                       \\
        \hline \hline
        Mozilla Firefox       & daneoshiga, bzum                    \\
        Evince                & hick209, bzum, marcicano, mquasar   \\
        BePDF / Xpdf          & marcicano                           \\
        Ekiga                 & flecha                              \\
        Empathy               & fefo                                \\
        Lib Folks (Telepathy) & kamiarc                             \\
        Scilab                & v1z, humannoise                     \\
        VxL                   & v1z                                 \\
        ImageMagick           & v1z                                 \\
        OpenOffice            & hick209                             \\
        Puredata              & v1z, automata, greenkobold, gilson, bzum \\
        Puredata OpenCV       & v1z                                 \\
        Puredata GEM          & v1z, fefo, hick209                  \\
        Puredata PDP          & v1z, fefo, hick209                  \\
        ChucK                 & rfabbri, automata                   \\
        ChucK MiniAudicle     & rfabbri, automata                   \\
        WebRTC                & automata                            \\
        OSC-Web               & automata                            \\
        Web-PD-GUI            & automata                            \\
        Live-Processing       & automata                            \\
        ChucK-Wiimote         & automata                            \\
        Audiolet              & automata                            \\
        Extempore             & automata                            \\
        \hline
    \end{tabular}
    \label{tabela:contribuicoes}
\end{table}
